---
title: "DATA607-Lab3B"
author: "Renan Charles-Pierre"
date: "2026-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(data.table)
library(slider)
```

## Introduction:

**Window Functions:\
**The data set that I have decided to use for this lab was a synthetic data set generated by ChatGPT. I prompted ChatGPT to generate a data set that included a time series for two or more items, I added the window functions and asked that the table be NBA themed because I wanted to base it around a topic that I find interesting. The generated data set was called NBA Daily Team Strength and features NBA team scores from 2022-2024. The set also includes a column called Team_Strength_Index which is a synthetic stat that is updated after every game. This column is the one that I will be using Window Functions on. I believe that the hardest part of the problem was figuring out what data set I could find that would fit the criteria, and now I believe the rest of the lab will be straightforward. I will update my conclusions if there are any changes.

```{r window-functions}
nba_daily_team_strength_df = read.csv("https://raw.githubusercontent.com/Renagade316/DATA607-Labs/refs/heads/main/Lab3/Lab3B/NBA%20Daily%20Team%20Strength%202022%20-%202024%20-%20Daily%20Team%20Data.csv")

nba_daily_team_strength_df <- select(nba_daily_team_strength_df, observation_date, calendar_year, team_name, opponent_abbreviation, is_game_day, win_flag, team_strength_index )

# Year to Date Average
nba_daily_team_strength_df <- nba_daily_team_strength_df |> 
  group_by(team_name) |> 
  mutate(year_to_date_avg = cummean(team_strength_index))


#6 Day Average
#slide_dbl(1:5, ~mean(.x), .before = 2)
nba_daily_team_strength_df <- nba_daily_team_strength_df |>
  group_by(team_name) |>
  mutate(six_day_avg = slide_mean(team_strength_index, before=5))

ex <- select(nba_daily_team_strength_df, team_name, team_strength_index, year_to_date_avg, six_day_avg)
ex <- filter(ex, team_name=="Boston Celtics")
ex

```

## Conclusions:

In conclusion, as I stated in my introduction, the project was relatively straightforward after getting my data set. I did have to spend extra time researching Window Functions. I now understand that Window Functions are useful for keeping track of data as it appears. Problems that I did have trouble with, was if the before in slide_mean was inclusive or not. To test my code, I created a data frame called *ex*. I only selected the team name, the team strength index, the year to date average and the six day average as these were the only values necessary for this purpose. Originally I had written, "six_day_avg = slide_mean(team_strength_index, before=6", but when I ran the code, I noticed that the 7th row the year to date average and six day average were the same number, this meant that the average for six days include the first date in the table. I double checked the average with my calculator to ensure that it was in fact calculating the average, which confirmed that the issue was with my before clause. I then switched the clause with before = 5, and ran the code again. This time the numbers were different, and the 6 day average ws from the 2nd to 7th date respectively.\
\
Link to ChatGPT prompt: <https://chatgpt.com/s/t_6aadef467bf08191ad9bdbcd35a5b6cb>

# Citations

OpenAI. (2026). *Generate an NBA time-series dataset for window-function analysis* [ChatGPT conversation]. ChatGPT.
