---
title: "DATA607-Lab3A"
author: "Renan Charles-Pierre"
date: "2026-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction:

**Window Functions:\
1.** Find (or ask an LLM to generate!) a dataset that includes time series for two or more separate items.  For example, you could use end of day stock or cryptocurrency prices since Jan 1, 2022 for several instruments.

2.  Use window functions (in SQL or dplyr) to calculate the year-to-date average and the six-day moving averages for each item.

The data set that I have decided to use for this lab was a synthetic data set generated by ChatGPT. I prompted ChatGPT to generate a data set that included a time series for two or more items, I added the window functions and asked that the table be NBA themed because I wanted to base it around a topic that I find interesting. The generated data set was called NBA Daily Team Strength and features NBA team scores from 2022-2024. The set also includes a column called Team_Strength_Index which is a synthetic stat that is updated after every game. This column is the one that I will be using Window Functions on. I believe that the hardest part of the problem was figuring out what data set I could find that would fit the criteria, and now I believe the rest of the lab will be straightforward. I will update my conclusions if there are any changes.

```{r window-functions}

```

## Conclusions:
